### 2021-10-21 submission/manuscript.tex
Show raw data that went into correlations

\textbf{The nature of singletons.} Removal of rare ASVs is commonly
justified as a method of removing ASVs that are artifacts. If such ASVs
are artifacts, then one would expect the number of singleton ASVs to
accumulate with sequencing depth. Contrary to this expectation, the
median percentage of sequences that were discarded when singleton ASVs
were removed from each dataset varied between 0.42 and 22.23\%
(bioethanol and seagrass). In addition, with the exception of the
samples from the marine and sediment datasets (Spearman correlation,
P\textgreater0.05), the fraction of singleton ASVs in samples was
negatively correlated with the number of sequences in each sample with a
range between -0.27 and -0.87 (rice and bioethanol) (Figure 1A). This
showed that with additional sequencing, the probability of seeing
singleton ASVs in multiple samples was greater than the probability of
generating an artifact. This suggests that the singleton ASVs are not as
likely to be artifacts as previously thought. Furthermore, if singleton
ASVs were artifacts, then one would not expect to find them in other
samples from the same dataset. In fact, singleton ASVs from samples with
fewer sequences were often found in samples with more sequences. At
least 50\% of the singleton ASVs found in the samples from the mice,
rice, seagrass, and stream datasets were found in another sample from
the same dataset (Figure 1B). Considering the likelihood of finding an
ASV duplicated in another sample is confounded by the number of samples
and inter-sample diversity, the high coverage of singleton ASVs in these
datasets was remarkable. The correlation between the number of sequences
in a sample and the fraction of that sample's singleton ASVs that were
covered by another sample in the dataset was significant and negative
for 9 of the datasets ranging between -0.31 and -0.84 for the rice and
seagrass datasets, respectively (Figure 1C). The negative correlation
indicated that the singleton ASVs in the smaller samples were more
likely to be covered by ASVs in the larger samples. Among the three
datasets without a significant correlation (Spearman correlation,
P\textgreater0.05), the marine and soil datasets had the fewest samples
in our collection and the stream dataset already had a high level of
coverage regardless of the number of sequences. Contrary to the common
motivation for removing rare ASVs, these results indicate that this
practice disproportionately impacts samples with fewer sequences and
likely removes more non-artifact ASVs than those that are artifacts.

\textbf{The impact of removing rare ASVs on the information represented
in each sample.} Removing rare ASVs will reduce the richness of ASVs
(i.e.~the number of ASVs per sample) and increase the relative abundance
of the remaining ASVs. To quantify the effect of removing rare ASVs on
the information contained within each sample, I varied the minimum
abundance threshold to simulate removing ASVs of varying rarity from
each sample. The richness of ASVs in each sample decreased by between
34.4 and 86.2\% (peromyscus and soil) when removing those ASVs that only
appeared once and by between 76.0 and 95.6\% (sediment and soil) when
removing those that appeared ten or fewer times from each sample (Figure
2A). Similarly, the Shannon diversity decreased by between 1.8 and
15.9\% (human and soil) when removing ASVs that only appeared once and
by between 5.4 and 35.4\% (human and seagrass) when removing ASVs that
appeared ten or fewer times from each sample (Figure 2B). Next, I
assigned the ASVs to OTUs, which were defined as a group of ASVs that
were more than 97\% similar to each other to assess the impact of
removing rare ASVs on higher level taxonomic groupings that are commonly
used in microbial ecology studies. Although pooling similar ASVs into
OTUs reduced the impact of removing the rare ASVs relative to the
ASV-based analysis, the minimum abundance threshold still decreased the
richness of OTUs and the diversity decreased relative to the full
community (Figure S2AB). In contrast to the richness and diversity
measurements, the Kullback--Leibler divergence compares the relative
abundance of specific ASVs or OTUs between representations of the
community. I calculated the Kullback--Leibler divergence between the
full communities and those where rare ASVs were removed. As the
threshold for removing ASVs increased, the amount of information lost
also increased for both ASVs and OTUs (Figure 2C and Figure S2C). The
relative loss of information was generally smaller for OTUs than than it
was for ASVs. Removing rare ASVs, regardless of abundance threshold, had
profound impacts on the representation of the communities.

\textbf{Removing treatment group effects from community data.} Because
treatment effects often affect a sample's diversity and inter-sample
variation, I generated null distributions for each study by randomizing,
without replacement, the number of times each ASV was observed in each
sample such that the total number of sequences in each sample and the
total number of times each ASV was observed across all samples in the
study was the same as was originally observed. This effectively made
every community in a study a statistical sample of the study-wide
composite community distribution. For example, after this procedure,
each of the 490 samples from the human dataset would be expected to have
the same richness and diversity of ASVs and one would not expect to find
treatment-based effects between the samples. Because of the risk of bias
if only one representation of the null distribution was generated, I
generated 100 randomized datasets for each study. The trends between
removing rare ASVs and the richness, diversity, and information loss
that were identified using the observed community distribution data were
also identified with the data from the null distribution; however, the
losses were larger when using the null distribution data (Figure S3).
The null distribution data were used in the remainder of the study to
minimize the risk of bias.

\textbf{The impact of removing rare ASVs on the information represented
between samples.} Considering the loss of richness, diversity, and
information when a community has its rarest ASVs removed, it seemed
likely that the relationship between communities would also be altered.
To assess the impact of removing rare ASVs on measures of alpha
diversity between samples I calculated the coefficients of variation
(COVs, i.e.~the standard deviation divided by the mean) for richness and
diversity for each study at multiple abundance thresholds. The COVs for
the richness of ASVs across the studies after removing singletons were
between 3.6 and 32.7-times larger than they were without removing
singleton ASVs (mice and stream; Figure 3A). Similarly, the COVs for the
diversity of ASVs were between 1.8 and 20.4-times larger when singletons
were removed than when they were not removed (mice and rice; Figure 3B).
To assess the impact of removing rare ASVs on measures of beta diversity
between samples, I calculated the COVs of the Bray-Curtis distances
between samples within the same study at multiple abundance thresholds.
The COVs between Bray-Curtis distances within a study when singletons
were removed was between 1.3 and 18.6-times larger than when they were
not removed (mice and stream; Figure 3C). When ASVs were clustered into
OTUs the difference in COVs was less than it was for the ASVs (Figure
S4). These results indicate that removing rare ASVs increases the
dissimilarity between samples, which could have a significant impact on
the statistical power to detect differences between treatment groups.

\textbf{The impact of removing rare ASVs on the ability to detect
statistically significant differences between treatment groups.} To test
the effect of increased inter-sample variation, I randomly assigned
samples to one of two treatment groups. In the first treatment group,
communities were randomly sampled from the null distribution as
described above. For the second treatment group, I randomly selected
10\% of the ASVs in the pooled study distribution to increase their
abundance by 5\%. I randomly generated 100 simulated sets of treatment
groups and samples. I then tested the ability to detect a difference
between the two treatment groups using alpha and beta diversity metrics.
The fraction of significant tests was a measurement of the statistical
power to detect the difference between the treatment groups. When
considering the differences in richness and diversity, the marine
dataset yielded no simulated sets that were statistically significant,
which was likely due to the small number of samples in the study (N=7).
Among the remaining datasets, the power to detect a difference in the
richness of ASVs ranged between 0.10 and 0.49 (sediment and stream) and
between 0.10 and 0.53 (rainforest and stream) to detect a difference in
diversity when using a Wilcox test (Figure 4A). When singleton ASVs were
removed, the power to detect a difference in the richness of ASVs
dropped by between 27.3 and 92.9\% (bioethanol and soil) and by between
40.0 and 93.3\% for their diversity(rainforest and soil; Figure 4B). The
effect of removing rare ASVs on the richness of OTUs and their diversity
was similar (Figure S5AB). I used the Bray-Curtis dissimilarity index to
compare the simulated communities within each dataset and calculated the
power to detect differences between the two simulated treatment groups
using the analysis of molecular variance (Figure 4C and S5C). Without
removing rare sequences, the power to detect a difference between the
two simulated treatment groups varied between 0.41 and 1.00 (rainforest
and rainforest). Aside from the bioethanol, human, and mice datasets,
the power to detect differences dropped by between 6.5 and 64.0\% (soil
and rice) when singletons were removed. However, when ASVs that occurred
10 or fewer times were removed from each sample, the power to detect
differences dropped by 12.0 and 97.2\% (human and peromyscus); similar
results were observed when ASVs were clustered into OTUs. Removing rare
ASVs reduced the ability to detect simulated treatment effects using
metrics commonly used to compare microbial communities.

\textbf{The impact of removing rare ASVs on the probability of falsely
detecting a difference between treatment groups.} I next asked whether
removing rare ASVs could lead to falsely claiming that a treatment
effect had a significant effect on community diversity and structure.
First, I sampled sequences from the null distribution for each dataset
and randomly assigned each sample to one of two treatment groups and
determined the richness and diversity of ASVs and OTUs. Testing at an
experiment-wise error rate of 0.05, I expected 5\% of the iterations for
each dataset to yield a significant test result. Indeed, there was no
evidence that removing rare ASVs resulted in an inflated experiment-wise
error rate. The average fraction of significant tests did not
meaningfully vary from 0.05 across the minimum abundance threshold,
dataset, metric of describing sample alpha-diversity, or whether the
abundance of ASVs or OTUs were used (Figure 5A and S6A). Similarly, the
average fraction of significant tests did not meaningfully vary from
0.05 when using analysis of molecular variance to compare communities
using Bray-Curtis distances (Figure 5A and S6A). Second, I again sampled
sequences from the null distribution, but assigned samples to one of two
treatment groups based on the number of sequences in each sample. The
samples with fewer than the median number of sequences for the dataset
were assigned to one group and those with more than the median were
assigned to the other. This exaggerated bias has been observed in
comparisons of the lung and oral microbiota because of the larger number
of non-specific amplicons that can be sequenced from lung samples
relative to those in the oral cavity leading to a significant difference
in sequencing depth between treatment groups (9). When rare sequences
were not removed, the fraction of significant tests did not differ from
5\% for comparing the richness, their diversity, or Bray-Curtis
distances (Figure 5B and S6B). However, when rare taxa of any frequency
were removed, the probability of falsely detecing a difference as
signifiant increased with the definition of rarity (Figure 5B and S6B).
Not including the small marine dataset, the average fraction the average
fraction of falsely detecting a difference across datasets when only
singletons were removed was 92.45\%. If there is any relationship
between the number of sequences and the treatment group, the risk of
falsely rejecting the null hypothesis is inflated when researchers use
the strategy of removing rare sequences. The most conservative approach
is to not remove low abundance sequences.

@@

Even with the best sequencing approaches, PCR artifacts and sequencing
errors persist. This may account for some sequences not being observed
other samples. Although this lack of inter-sample coverage could have
been due to treatment effects and natural variation, it is likely that
some portion of the sequences that were unique to samples were artifacts
or contained errors. Researchers are encouraged to use rarefaction to
control for uneven sampling and the presence of spurious sequences.
Previous work sequencing mock communities has shown that the number of
spurious sequences increases with sequencing depth (2, 14). By rarefying
data to a common number of sequences per sample, the number of spurious
sequences can be controlled. As shown in the data I presented, which was
rarefied to a common number of reads per sample within a dataset, when
rare sequences were not removed the power to detect differences was the
highest and the false discovery rate was the expected 5\% (Figures 5 and
S6).

@@

\textbf{Figure 1. Singletons are more common in samples with fewer
seqeunces and tend to be shared with samples having more sequences.} For
each of the 12 datasets, Spearman correlation coefficients were
calcualted between the number of sequences in each sample and the number
of singletons in the sample (A) and the fraction of its singletons that
were shared with another sample (C). Those correlations that were not
statisically significant had a P-value greater than 0.05. The faction of
singletons shared across samples (B) were calculated for each dataset.
The median value is shown with a solid circle and the 95\% confidence
interval is indicated by the solid line.

@@

\textbf{Figure S2. Removing rare sequences from samples alters their
representation of alpha-diversity using operational taxonomic units
(OTUs).} The average difference in the richness (A), Shannon diversity
(B), and Kullback-Leiber divergence (C) for each sample within a dataset
was calculated between the original community structures relative to
applying different minimum abundance thresholds.

@@

\textbf{Figure S3. Removing rare sequences from samples alters their
representation of alpha-diversity when regenerating samples using a null
distribution for each dataset.} The average difference in the richness,
Shannon diversity, and Kullback-Leiber divergence for each sample within
a dataset was calculated between the original community structures
relative to applying different minimum abundance thresholds. Some values
of Kullback-Leiber divergence are missing because undefined values were
calculated due to the removal of rare sequences. Data are shown for
amplicon sequence variants (ASVs) and operational taxonomic units (OTUs)

@@

\textbf{Figure S4. Removing rare sequences from samples increases the
inter-sample variation for operational taxonomic units (OTUs).} The
coefficient of variation in richness (A), Shannon diversity (B), and
Bray-Curtis distances (C) for each dataset was calculated using the null
distributed samples for each dataset with varying minimum abundance
thresholds.

@@

\textbf{Figure S5. Removing rare sequences from samples reduces the
statistical power to detect differences between empirically generated
treatment groups when using operational taxonomic units (OTUs).} The
fraction of significant tests comparing the richness (A) and Shannon
diversity (B) using a Wilcox test and Bray-Curtis distances (C) using
analysis of molecular variance for each dataset was calculated using
empirically generated treatment groups containing equal numbers of
samples for each dataset with varying minimum abundance thresholds. For
each dataset and minimum abundance threshold, 100 randomizations were
peformed.

@@

\textbf{Figure S6. Removing rare sequences does not impact the false
detection rate unless the number of sequences per sample is confounded
with the treatment groups when using operational taxonomic units
(OTUs).} The fraction of significant tests comparing the richness and
Shannon diversity using a Wilcox test and Bray-Curtis distances using
analysis of molecular variance for each dataset was calculated.
Empirically generated treatment groups were generated containing equal
numbers of samples where the samples represented a null distribution. In
one simulation the samples were randomly assigned to a treatment group
(A) and in the other the samples were assigned based on the number of
sequences in each sample (B). For each dataset and minimum abundance
threshold, 100 randomizations were peformed.

>>>>

\textbf{The nature of singletons.} Removal of rare ASVs is commonly
justified as a method of removing ASVs that are artifacts. If such ASVs
are artifacts, then one would expect the number of singleton ASVs to
accumulate with sequencing depth. Contrary to this expectation, the
median percentage of sequences that were discarded when singleton ASVs
were removed from each dataset varied between 0.42 and 22.20\%
(bioethanol and seagrass). In addition, with the exception of the
samples from the marine and sediment datasets (Spearman correlation,
P\textgreater0.05), the fraction of singleton ASVs in samples was
negatively correlated with the number of sequences in each sample with a
range between -0.26 and -0.87 (rice and bioethanol) (Figures 1A and S2).
This showed that with additional sequencing, the probability of seeing
singleton ASVs in multiple samples was greater than the probability of
generating an artifact. This suggests that the singleton ASVs are not as
likely to be artifacts as previously thought. Furthermore, if singleton
ASVs were artifacts, then one would not expect to find them in other
samples from the same dataset. In fact, singleton ASVs from samples with
fewer sequences were often found in samples with more sequences. At
least 50\% of the singleton ASVs found in the samples from the mice,
rice, seagrass, and stream datasets were found in another sample from
the same dataset (Figure 1B). Considering the likelihood of finding an
ASV duplicated in another sample is confounded by the number of samples
and inter-sample diversity, the high coverage of singleton ASVs in these
datasets was remarkable. The correlation between the number of sequences
in a sample and the fraction of that sample's singleton ASVs that were
covered by another sample in the dataset was significant and negative
for 9 of the datasets ranging between -0.33 and -0.84 for the rice and
seagrass datasets, respectively ((Figures 1C and S3)). The negative
correlation indicated that the singleton ASVs in the smaller samples
were more likely to be covered by ASVs in the larger samples. Among the
three datasets without a significant correlation (Spearman correlation,
P\textgreater0.05), the marine and soil datasets had the fewest samples
in our collection and the stream dataset already had a high level of
coverage regardless of the number of sequences. Contrary to the common
motivation for removing rare ASVs, these results indicate that this
practice disproportionately impacts samples with fewer sequences and
likely removes more non-artifact ASVs than those that are artifacts.

\textbf{The impact of removing rare ASVs on the information represented
in each sample.} Removing rare ASVs will reduce the richness of ASVs
(i.e.~the number of ASVs per sample) and increase the relative abundance
of the remaining ASVs. To quantify the effect of removing rare ASVs on
the information contained within each sample, I varied the minimum
abundance threshold to simulate removing ASVs of varying rarity from
each sample. The richness of ASVs in each sample decreased by between
34.4 and 86.2\% (peromyscus and soil) when removing those ASVs that only
appeared once and by between 76.0 and 95.6\% (sediment and soil) when
removing those that appeared ten or fewer times from each sample (Figure
2A). Similarly, the Shannon diversity decreased by between 1.8 and
15.9\% (human and soil) when removing ASVs that only appeared once and
by between 5.4 and 35.4\% (human and seagrass) when removing ASVs that
appeared ten or fewer times from each sample (Figure 2B). Next, I
assigned the ASVs to OTUs, which were defined as a group of ASVs that
were more than 97\% similar to each other to assess the impact of
removing rare ASVs on higher level taxonomic groupings that are commonly
used in microbial ecology studies. Although pooling similar ASVs into
OTUs reduced the impact of removing the rare ASVs relative to the
ASV-based analysis, the minimum abundance threshold still decreased the
richness of OTUs and the diversity decreased relative to the full
community (Figure S4AB). In contrast to the richness and diversity
measurements, the Kullback--Leibler divergence compares the relative
abundance of specific ASVs or OTUs between representations of the
community. I calculated the Kullback--Leibler divergence between the
full communities and those where rare ASVs were removed. As the
threshold for removing ASVs increased, the amount of information lost
also increased for both ASVs and OTUs (Figure 2C and Figure S4C). The
relative loss of information was generally smaller for OTUs than than it
was for ASVs. Removing rare ASVs, regardless of abundance threshold, had
profound impacts on the representation of the communities.

\textbf{Removing treatment group effects from community data.} Because
treatment effects often affect a sample's diversity and inter-sample
variation, I generated null distributions for each study by randomizing,
without replacement, the number of times each ASV was observed in each
sample such that the total number of sequences in each sample and the
total number of times each ASV was observed across all samples in the
study was the same as was originally observed. This effectively made
every community in a study a statistical sample of the study-wide
composite community distribution. For example, after this procedure,
each of the 490 samples from the human dataset would be expected to have
the same richness and diversity of ASVs and one would not expect to find
treatment-based effects between the samples. Because of the risk of bias
if only one representation of the null distribution was generated, I
generated 100 randomized datasets for each study. The trends between
removing rare ASVs and the richness, diversity, and information loss
that were identified using the observed community distribution data were
also identified with the data from the null distribution; however, the
losses were larger when using the null distribution data (Figure S5).
The null distribution data were used in the remainder of the study to
minimize the risk of bias.

\textbf{The impact of removing rare ASVs on the information represented
between samples.} Considering the loss of richness, diversity, and
information when a community has its rarest ASVs removed, it seemed
likely that the relationship between communities would also be altered.
To assess the impact of removing rare ASVs on measures of alpha
diversity between samples I calculated the coefficients of variation
(COVs, i.e.~the standard deviation divided by the mean) for richness and
diversity for each study at multiple abundance thresholds. The COVs for
the richness of ASVs across the studies after removing singletons were
between 3.6 and 32.7-times larger than they were without removing
singleton ASVs (mice and stream; Figure 3A). Similarly, the COVs for the
diversity of ASVs were between 1.8 and 20.4-times larger when singletons
were removed than when they were not removed (mice and rice; Figure 3B).
To assess the impact of removing rare ASVs on measures of beta diversity
between samples, I calculated the COVs of the Bray-Curtis distances
between samples within the same study at multiple abundance thresholds.
The COVs between Bray-Curtis distances within a study when singletons
were removed was between 1.3 and 18.6-times larger than when they were
not removed (mice and stream; Figure 3C). When ASVs were clustered into
OTUs the difference in COVs was less than it was for the ASVs (Figure
S6). These results indicate that removing rare ASVs increases the
dissimilarity between samples, which could have a significant impact on
the statistical power to detect differences between treatment groups.

\textbf{The impact of removing rare ASVs on the ability to detect
statistically significant differences between treatment groups.} To test
the effect of increased inter-sample variation, I randomly assigned
samples to one of two treatment groups. In the first treatment group,
communities were randomly sampled from the null distribution as
described above. For the second treatment group, I randomly selected
10\% of the ASVs in the pooled study distribution to increase their
abundance by 5\%. I randomly generated 100 simulated sets of treatment
groups and samples. I then tested the ability to detect a difference
between the two treatment groups using alpha and beta diversity metrics.
The fraction of significant tests was a measurement of the statistical
power to detect the difference between the treatment groups. When
considering the differences in richness and diversity, the marine
dataset yielded no simulated sets that were statistically significant,
which was likely due to the small number of samples in the study (N=7).
Among the remaining datasets, the power to detect a difference in the
richness of ASVs ranged between 0.10 and 0.49 (sediment and stream) and
between 0.10 and 0.53 (rainforest and stream) to detect a difference in
diversity when using a Wilcox test (Figure 4A). When singleton ASVs were
removed, the power to detect a difference in the richness of ASVs
dropped by between 27.3 and 92.9\% (bioethanol and soil) and by between
40.0 and 93.3\% for their diversity(rainforest and soil; Figure 4B). The
effect of removing rare ASVs on the richness of OTUs and their diversity
was similar (Figure S7AB). I used the Bray-Curtis dissimilarity index to
compare the simulated communities within each dataset and calculated the
power to detect differences between the two simulated treatment groups
using the analysis of molecular variance (Figure 4C and S7C). Without
removing rare sequences, the power to detect a difference between the
two simulated treatment groups varied between 0.41 and 1.00 (rainforest
and rainforest). Aside from the bioethanol, human, and mice datasets,
the power to detect differences dropped by between 6.5 and 64.0\% (soil
and rice) when singletons were removed. However, when ASVs that occurred
10 or fewer times were removed from each sample, the power to detect
differences dropped by 12.0 and 97.2\% (human and peromyscus); similar
results were observed when ASVs were clustered into OTUs. Removing rare
ASVs reduced the ability to detect simulated treatment effects using
metrics commonly used to compare microbial communities.

\textbf{The impact of removing rare ASVs on the probability of falsely
detecting a difference between treatment groups.} I next asked whether
removing rare ASVs could lead to falsely claiming that a treatment
effect had a significant effect on community diversity and structure.
First, I sampled sequences from the null distribution for each dataset
and randomly assigned each sample to one of two treatment groups and
determined the richness and diversity of ASVs and OTUs. Testing at an
experiment-wise error rate of 0.05, I expected 5\% of the iterations for
each dataset to yield a significant test result. Indeed, there was no
evidence that removing rare ASVs resulted in an inflated experiment-wise
error rate. The average fraction of significant tests did not
meaningfully vary from 0.05 across the minimum abundance threshold,
dataset, metric of describing sample alpha-diversity, or whether the
abundance of ASVs or OTUs were used (Figure 5A and S8A). Similarly, the
average fraction of significant tests did not meaningfully vary from
0.05 when using analysis of molecular variance to compare communities
using Bray-Curtis distances (Figure 5A and S8A). Second, I again sampled
sequences from the null distribution, but assigned samples to one of two
treatment groups based on the number of sequences in each sample. The
samples with fewer than the median number of sequences for the dataset
were assigned to one group and those with more than the median were
assigned to the other. This exaggerated bias has been observed in
comparisons of the lung and oral microbiota because of the larger number
of non-specific amplicons that can be sequenced from lung samples
relative to those in the oral cavity leading to a significant difference
in sequencing depth between treatment groups (9). When rare sequences
were not removed, the fraction of significant tests did not differ from
5\% for comparing the richness, their diversity, or Bray-Curtis
distances (Figure 5B and S8B). However, when rare taxa of any frequency
were removed, the probability of falsely detecing a difference as
signifiant increased with the definition of rarity (Figure 5B and S8B).
Not including the small marine dataset, the average fraction the average
fraction of falsely detecting a difference across datasets when only
singletons were removed was 92.45\%. If there is any relationship
between the number of sequences and the treatment group, the risk of
falsely rejecting the null hypothesis is inflated when researchers use
the strategy of removing rare sequences. The most conservative approach
is to not remove low abundance sequences.

@@

Even with the best sequencing approaches, PCR artifacts and sequencing
errors persist. This may account for some sequences not being observed
other samples. Although this lack of inter-sample coverage could have
been due to treatment effects and natural variation, it is likely that
some portion of the sequences that were unique to samples were artifacts
or contained errors. Researchers are encouraged to use rarefaction to
control for uneven sampling and the presence of spurious sequences.
Previous work sequencing mock communities has shown that the number of
spurious sequences increases with sequencing depth (2, 14). By rarefying
data to a common number of sequences per sample, the number of spurious
sequences can be controlled. As shown in the data I presented, which was
rarefied to a common number of reads per sample within a dataset, when
rare sequences were not removed the power to detect differences was the
highest and the false discovery rate was the expected 5\% (Figures 5 and
S8).

@@

\textbf{Figure 1. Singletons are more common in samples with fewer
seqeunces and tend to be shared with samples having more sequences.} For
each of the 12 datasets, Spearman correlation coefficients were
calcualted between the number of sequences in each sample and the number
of singletons in the sample (A) and the fraction of its singletons that
were shared with another sample (C). Those correlations that were not
statisically significant had a P-value greater than 0.05. The faction of
singletons shared across samples (B) were calculated for each dataset.
The median value is shown with a solid circle and the 95\% confidence
interval is indicated by the solid line. The raw data for A and C are
presented in Figures S2 and S3, respectively.

@@

\textbf{Figure S2. Singletons are more common in samples with fewer
seqeunces.} Spearman correlation coefficients (\(\rho\)) were calcualted
between the total number of sequences and the number of singletons in
the sample. The Spearman correlation coefficients for the Marine and
Sediment datasets were not significantly different from 0.00
(P\textgreater0.05). The blue line indicates the linear trend in the
data. These data are summarized in Figure 1A.

@@

\textbf{Figure S3. Singletons tend to be shared with samples having more
sequences.} Spearman correlation coefficients (\(\rho\)) were calcualted
between the total number of sequences and the number of singletons in a
sample that were shared with other samples from the same deataset. The
Spearman correlation coefficients for the Marine, Soil, and Stream
datasets were not significantly different from 0.00 (P\textgreater0.05).
The blue line indicates the linear trend in the data. These data are
summarized in Figure 1C.

@@

\textbf{Figure S4. Removing rare sequences from samples alters their
representation of alpha-diversity using operational taxonomic units
(OTUs).} The average difference in the richness (A), Shannon diversity
(B), and Kullback-Leiber divergence (C) for each sample within a dataset
was calculated between the original community structures relative to
applying different minimum abundance thresholds.

@@

\textbf{Figure S5. Removing rare sequences from samples alters their
representation of alpha-diversity when regenerating samples using a null
distribution for each dataset.} The average difference in the richness,
Shannon diversity, and Kullback-Leiber divergence for each sample within
a dataset was calculated between the original community structures
relative to applying different minimum abundance thresholds. Some values
of Kullback-Leiber divergence are missing because undefined values were
calculated due to the removal of rare sequences. Data are shown for
amplicon sequence variants (ASVs) and operational taxonomic units (OTUs)

@@

\textbf{Figure S6. Removing rare sequences from samples increases the
inter-sample variation for operational taxonomic units (OTUs).} The
coefficient of variation in richness (A), Shannon diversity (B), and
Bray-Curtis distances (C) for each dataset was calculated using the null
distributed samples for each dataset with varying minimum abundance
thresholds.

\includegraphics{figure_s7.png}

\textbf{Figure S7. Removing rare sequences from samples reduces the
statistical power to detect differences between empirically generated
treatment groups when using operational taxonomic units (OTUs).} The
fraction of significant tests comparing the richness (A) and Shannon
diversity (B) using a Wilcox test and Bray-Curtis distances (C) using
analysis of molecular variance for each dataset was calculated using
empirically generated treatment groups containing equal numbers of
samples for each dataset with varying minimum abundance thresholds. For
each dataset and minimum abundance threshold, 100 randomizations were
peformed.

\includegraphics{figure_s8.png}

\textbf{Figure S8. Removing rare sequences does not impact the false
detection rate unless the number of sequences per sample is confounded
with the treatment groups when using operational taxonomic units
(OTUs).} The fraction of significant tests comparing the richness and
Shannon diversity using a Wilcox test and Bray-Curtis distances using
analysis of molecular variance for each dataset was calculated.
Empirically generated treatment groups were generated containing equal
numbers of samples where the samples represented a null distribution. In
one simulation the samples were randomly assigned to a treatment group
(A) and in the other the samples were assigned based on the number of
sequences in each sample (B). For each dataset and minimum abundance
threshold, 100 randomizations were peformed.
